---
title: "Environmental Impact of T-shirt Production"
author: "Emily Wilson"
date: "6-2-2026"
format:
  html:
    theme: cosmo
editor_options: 
  chunk_output_type: inline
---
# Intro 
My story is examining the environmental impact of different materials used to make T-shirts — one of the most common promotional products distributed in the United States. The environmental impact is measured by carbon footprint and water usage, and the materials I’ll be looking at are five different types of cotton blends. To put the environmental impact into context, I’ll work in data from Advertising Specialty Institute’s 2026 end-user survey about impressions and longevity, and the company’s calculation of impressions over lifespan for T-shirts (3,500).  

The pitch has been discussed with my editors at Advertising Specialty Institute and largely green-lit. The company’s audience is the promotional products industry, so the intention with this story is to give readers more information on the types of t-shirts they’re buying. There are lots of factors to consider — price, quality, etc — but more organized information about the different materials used to make t-shirts and its respective environmental impact might be helpful for potential buyers to consider. 

The bulk of my work over the course of this week will be to source data on the carbon and water usage from the different cotton blends and document that process. I’ve also created an empty table to help visualize the data I’ll need to find. Rows 4 and 5 are calculations I'll complete once rows 2 and 3 are done. 

```{r}
#| include: false
library(tidyverse)
library(readxl)
library(googlesheets4)
library(googledrive)
library(janitor)
```
```{r}
#| include: false
asi_data_needed <- rio::import("https://docs.google.com/spreadsheets/d/1HyKYvl8NUcbQiNJRf6-M9yBzVGHWB8zAVFOjiHMVTeI/edit?gid=0#gid=0") 
```

```{r}
#| include: false
asi_data_needed <- rio::import("https://docs.google.com/spreadsheets/d/1HyKYvl8NUcbQiNJRf6-M9yBzVGHWB8zAVFOjiHMVTeI/edit?gid=0#gid=0") 
```

```{r}
#| echo: false
asi_data_needed
```

# Progress 
## ASI 2026 end-user survey data: 
This is where the 3,500 impressions per lifespan for t-shirts come from. The metric was calculated by the newsroom from the impressions and longevity data, and I found it in this report. 
I can also use this data for context on how often consumers wear their promo t-shirts and how many people see them when they do. In R, I’ve filtered down the dataset to only show this data. 
Ex: Around 80% of consumers reported that they kept their T-shirts for at least a year. 

```{r}
#| include: false
us_ads <- rio::import("https://github.com/djnf-data-2026/emily-wilson/raw/refs/heads/main/data/2026%20Ad%20Impressions%20US%20Products.xlsx")
```


```{r}
#| include: false
impressions <- us_ads[95:108, ]

colnames(impressions) <- impressions[1, ]

impressions <- impressions |>
  clean_names() |>
  as_tibble()

impressions <- impressions[-1, ] |>
  mutate(across(2:21, as.numeric)) |>
  rename(item = na)

impressions |> 
  select(1:21) |> 
  head()
```
```{r}
#| echo: false
tshirt_impressions <- impressions %>% select("item", "t_shirts")

tshirt_impressions
```
```{r}
#| include: false
longevity <- us_ads[118:131, ]

colnames(longevity) <- longevity[1, ]

longevity <- longevity |>
  clean_names() |>
  as_tibble()

longevity <- longevity[-1, ] |>
  mutate(across(2:21, as.numeric)) |>
  rename(item = na)
```
```{r}
#| include: false
tshirt_longevity <- longevity %>% select("item", "t_shirts")
```
```{r}
#| echo: false
tshirt_longevity
```


## Note: 
My editor at Advertising Specialty Institute, Theresa Hegel, had initially provided some different calculators for environmental impact. The majority of those are not usable for this analysis, but helpful context for understanding the different steps in the supply chain that can have an environmental impact on the T-shirts. 

## Bonsai:  
[Bonsai](https://www.bonsai.uno/?mode=product&metric=GWP100&version=v2.1.6&product=C_Cotton&region=US&advanced=1&unit=kg) is a company that produces sustainability information about products. They have data from cotton lint, ginned, which produces -2.897 kg of carbon dioxide per 1 kg produced. This is interesting information because it indicates that at least one step of cotton production can potentially have a positive impact on the environment by removing carbon emissions, but since it’s for cotton overall and not stratified by cotton type, it’s not necessarily something I can use. It’s also using data from 2016, which is not ideal. 

## Company calculators: 
Some suppliers that consider their products sustainable create calculators to compare their environmental impact with that of another company. These are unusable in the data analysis because the newsroom doesn’t want to use a calculator created by a company because there might be bias. One example is Everywhere Apparel, which uses recycled cotton (one of my categories!). That company has information that indicates that for every T-shirt they produce, it saves 1,123 gallons of water, about one fifth of a pound of atmospheric carbon emissions and reduces landfill waste by 0.6 pounds, compared to a regular T-shirt. This is helpful because it indicates the unit of measurement that I could be looking for (gallons for water, pounds for atmospheric carbon emissions). There is also Allmade, which again, only has data for its own products. They do separate the environmental impact based on cotton type, so I can find example data for organic cotton and tri-blend. For recycled cotton, they only have heavyweight T-shirt examples, and I’m not sure if that can be used for this. 

## Life Cycle Assessment - Cotton
Another resource that was provided is a Life Cycle Assessment for cotton produced by Textile Exchange, a nonprofit that conducts research on sustainability in the supply chain. This was helpful because it provided a thorough overview of the different steps in the supply chain for cotton production and compares different countries. The assessment uses nitrous oxide emissions to measure the environmental impact, as opposed to water anc carbon, so I'm not able to use it for my project. 


